### 2022-03-18 script/paper 3.Rmd
Add files via upload

---
title: "Untitled"
author: "Jiayi Du"
date: "2022/3/17"
output: pdf_document
---

```{r}
gss<-read.csv("gss.csv",header=TRUE)
gss<-gss%>%select(age,feelings_life,sex,marital_status,education,hh_size,average_hours_worked,income_family,income_respondent)
```

```{r}
gss<-gss%>%na.omit(gss)
gss<-gss%>%filter(marital_status=="Married"|marital_status=="Single, never married")%>%filter(average_hours_worked!="Don't know")

```

```{r}
library(ggplot2)
ggplot(gss,aes(x=feelings_life,fill=income_family))+geom_histogram(position = position_dodge(0.75),binwidth = 1)+facet_grid(.~marital_status)
ggplot(gss,aes(x=feelings_life,fill=average_hours_worked))+geom_histogram(position = position_dodge(0.75),binwidth = 1)+facet_grid(.~marital_status)

```

```{r}
ggplot(aes(x = age, fill = marital_status), data = gss) +geom_histogram(binwidth = 0.2)+facet_wrap(~feelings_life)
ggplot(aes(x = feelings_life, fill = sex), data = gss) +geom_bar(position = "dodge")+facet_wrap(~marital_status)
```

```{r}

```