### 2022-03-18 output/paper 3.Rmd
first

---
title: "Canadian Family"
subtitle: "My subtitle if needed"
author: 
  - Alicia Yang
  - [Another author]
  - []
thanks: "Code and data are available at: https://github.com/dujiayi1/Factors-that-affect-happiness."
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references.bib
---

```{r}
gss<- read_csv(here::here(".Factors-that-affect-happiness/input/gss_family.csv"))
gss<-gss%>%select(age,feelings_life,sex,marital_status,education,hh_size,average_hours_worked,income_family,income_respondent)
```

```{r}
gss<-gss%>%na.omit(gss)
gss<-gss%>%filter(marital_status=="Married"|marital_status=="Single, never married")%>%filter(average_hours_worked!="Don't know")

```

```{r}
library(ggplot2)
ggplot(gss,aes(x=feelings_life,fill=income_family))+geom_histogram(position = position_dodge(0.75),binwidth = 1)+facet_grid(.~marital_status)
ggplot(gss,aes(x=feelings_life,fill=average_hours_worked))+geom_histogram(position = position_dodge(0.75),binwidth = 1)+facet_grid(.~marital_status)

```

```{r}
ggplot(aes(x = age, fill = marital_status), data = gss) +geom_histogram(binwidth = 0.2)+facet_wrap(~feelings_life)
ggplot(aes(x = feelings_life, fill = sex), data = gss) +geom_bar(position = "dodge")+facet_wrap(~marital_status)
```

```{r}

```